\textbf{SPECFEM3D}\\ \textbf{User Manual}

### 2010-11-24 doc/USER_MANUAL/manual_SPECFEM3D.tex
fixed a bug with URLs and parentheses

\let\myUrl\url
\renewcommand{\url}[1]{(\myUrl{#1})}

@@

The cover graphic of the manual was created
by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}. \\

@@

Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}. We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \url{cig-seismo@geodynamics.org}.

@@

In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \url{cig-seismo@geodynamics.org}.

@@

\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \url{www.simulia.com}

@@

\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \url{www.opendx.org}

@@

save ParaView \url{www.paraview.org}

@@

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \url{www.seismology.harvard.edu}.
It looks like this:

@@

\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC \url{www.llnl.gov/sac}.
Type

@@

In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
in the mesher's parameter file \texttt{Mesh\_Par\_file}
prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
You can then use AVS \url{www.avs.com} or OpenDX \url{www.opendx.org}
to visualize the mesh and MPI partition (slices).

% obsolete: only for global package...
%Use the serial code \texttt{combine\_AVS\_DX.f90} (type `\texttt{make
%combine\_AVS\_DX}' and then `\texttt{xcombine\_AVS\_DX}') to generate
%AVS \url{www.avs.com} output files (in AVS UCD format) or OpenDX \url{www.opendx.org}
%output files showing the mesh, the MPI partition (slices), the $\nchunks$
%chunks, the source and receiver location, etc.
%Use the AVS UCD files \texttt{AVS\_continent\_boundaries.inp} and \texttt{AVS\_plate\_boundaries.inp}
%or the OpenDX files \texttt{DX\_continent\_boundaries.dx} and \texttt{DX\_plate\_boundaries.dx}
%for reference.

@@

You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
using Paraview \url{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
This will create VTK files for each single partition. You can then use Paraview \url{www.paraview.org} to visualized these partitions.

@@

The \texttt{SPECFEM3D} code is running in near real-time to produce
animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \url{www.shakemovie.caltech.edu}.

@@

\begin{lyxcode}
mesh2vtu.pl~-i~file.mesh~-o~file.vtu
\end{lyxcode}
\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \url{http://www.vtk.org/}
run-time library for its execution. Therefore, make sure you have
them properly set in the script according to your system.
\end{enumerate}
\item \textbf{Copy over the source and receiver .vtk file}

@@

Finally, we can view the mesh in ParaView \url{www.paraview.org}.

@@

\begin{itemize}
\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
only the bounding box is shown.
\end{itemize}
\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
\textsf{Representation} and select \textsf{wireframe of surface} to
display it.
\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
script (either from the standard output or from \texttt{normal\_plane.txt}
file).
\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
lower and upper limits, or change the color scheme.
\item Now load in the source and receiver location file by \textsf{File}
$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
\item For more information about ParaView, see the ParaView Users Guide \url{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
\end{enumerate}
\end{enumerate}
%
\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[scale=0.6]{figures/3D-S-Kernel}
\par\end{centering}

@@

\begin{enumerate}
\item First the script directs the scheduler to save its own output and
output from \texttt{stdout} into ~\\
\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
where \texttt{\%J} is short-hand for the job-ID; it also tells the
scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
\item The script then creates a list of the nodes allocated to this job
by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
and parsing the output into a one-column list using the Perl script
\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
directories on these nodes (\texttt{\small /scratch/}~\\
\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
for temporary storage of the database files. The scratch directories
are created using \texttt{shmux}, a shell multiplexor that can execute
the same commands on many hosts in parallel. \texttt{shmux} is available
from Shmux \url{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
\item The next portion of the script launches the mesher, database generator and then the solver
using \texttt{run\_lsf.bash}.
\item The final portion of the script collects the seismograms and performs
clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
and \texttt{cleanmulti.pl}.
\end{enumerate}
\begin{lyxcode}
\#!/bin/bash~-v~

@@

rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
\end{lyxcode}
where the first command performs rotation on the SAC data obtained
through Seismogram Transfer Program (STP) \url{http://www.data.scec.org/STP/stp.html},
while the second command rotates the processed SEM synthetics.

@@

To report bugs or suggest improvements to the code, please send an
e-mail to the CIG Computational Seismology Mailing List \url{cig-seismo@geodynamics.org}
or Jeroen Tromp \url{jtromp-AT-princeton.edu}, and/or use our online
bug tracking system Roundup \url{www.geodynamics.org/roundup}.

@@

OpenDX \url{http://www.opendx.org} is open-source based on IBM Data
Explorer, AVS \url{http://www.avs.com} is a trademark of Advanced
Visualization Systems, and ParaView \url{http://www.paraview.org}
is an open-source visualization platform.

%The main developers of the \texttt{SPECFEM3D} source code are Dimitri
%Komatitsch, Jeroen Tromp, and Qinya Liu. The following individuals
%(listed in alphabetical order) have also contributed to the development
%or improvement of the source code: Min Chen, Vala Hj\"orleifsd\'ottir,
%Jes\'us Labarta, and Leif Strand. The following individuals (listed in alphabetical
%order) contributed to this manual: Min Chen, Vala Hj\"orleifsd\'ottir,
%Sue Kientz, Dimitri Komatitsch, Qinya Liu, Alessia Maggi,
%Carl Tape, and Jeroen Tromp. The manual's cover graphic was created
%by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \url{http://www.cacr.caltech.edu}.
%Older versions of the code were initially developed by Dimitri Komatitsch at Institut de Physique du Globe (France)
%and then by Dimitri Komatitsch and Jeroen Tromp at Harvard University (USA).

Please e-mail your feedback, questions, comments, and suggestions
to Jeroen Tromp \url{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \url{cig-seismo@geodynamics.org}.

>>>>

\newcommand{\urlwithparentheses}[1]{(\url{#1})}

@@

The cover graphic of the manual was created
by Santiago Lombeyda from Caltech's Center for Advanced Computing Research (CACR) \urlwithparentheses{http://www.cacr.caltech.edu}. \\

@@

Examples are provided in the SPECFEM3D package in the subdirectory \texttt{examples/}. We strongly encourage you to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.

@@

In each example subdirectory you will find a \texttt{README} file, which explains in a step-by-step tutorial the workflow for the example.
Please feel free to contribute your own example to this package by contacting the CIG Computational Seismology Mailing List  \urlwithparentheses{cig-seismo@geodynamics.org}.

@@

\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.} to save Abaqus FEA \urlwithparentheses{www.simulia.com}

@@

\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to save OpenDX \urlwithparentheses{www.opendx.org}

@@

save ParaView \urlwithparentheses{www.paraview.org}

@@

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like this:

@@

\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC \urlwithparentheses{www.llnl.gov/sac}.
Type

@@

In case you used the internal mesher \texttt{xmeshfem3D} to create and partition your mesh,
you can output mesh files in ABAQUS (.INP) and DX (.dx) format to visualize them. For this, you must set either the
flag \texttt{CREATE\_DX\_FILES} or \texttt{CREATE\_ABAQUS\_FILES} to \texttt{.true.}
in the mesher's parameter file \texttt{Mesh\_Par\_file}
prior to running the mesher (see Chapter~\ref{cha:Running-the-Mesher-Meshfem3D} for details).
You can then use AVS \urlwithparentheses{www.avs.com} or OpenDX \urlwithparentheses{www.opendx.org}
to visualize the mesh and MPI partition (slices).

@@

You have also the option to visualize the distributed databases produced by \texttt{xgenerate\_databases}
using Paraview \urlwithparentheses{www.paraview.org}. For this, you must set the flag \texttt{SAVE\_MESH\_FILES} to \texttt{.true.}
in the main parameter file \texttt{Par\_file} (see Chapter~\ref{cha:Main-Parameter} for details).
This will create VTK files for each single partition. You can then use Paraview \urlwithparentheses{www.paraview.org} to visualized these partitions.

@@

The \texttt{SPECFEM3D} code is running in near real-time to produce
animations of southern California earthquakes via the web; see Southern California ShakeMovie\textregistered \urlwithparentheses{www.shakemovie.caltech.edu}.

@@

\begin{lyxcode}
mesh2vtu.pl~-i~file.mesh~-o~file.vtu
\end{lyxcode}
\item Notice that this Perl script uses a program \texttt{mesh2vtu} in the
\texttt{utils/Visualization/Paraview/mesh2vtu} directory, which further uses the VTK \urlwithparentheses{http://www.vtk.org/}
run-time library for its execution. Therefore, make sure you have
them properly set in the script according to your system.
\end{enumerate}
\item \textbf{Copy over the source and receiver .vtk file}

@@

Finally, we can view the mesh in ParaView \urlwithparentheses{www.paraview.org}.

@@

\begin{itemize}
\item If the mesh file is of moderate size, it shows up on the screen; otherwise,
only the bounding box is shown.
\end{itemize}
\item Click \textsf{Display Tab} $\rightarrow$ \textsf{Display Style} $\rightarrow$
\textsf{Representation} and select \textsf{wireframe of surface} to
display it.
\item To create a cross-section of the volumetric mesh, choose \textsf{Filter}
$\rightarrow$ \textsf{cut}, and under \textsf{Parameters Tab}, choose
\textsf{Cut Function} $\rightarrow$ \textsf{plane}.
\item Fill in center and normal information given by the \texttt{global\_slice\_number.pl}
script (either from the standard output or from \texttt{normal\_plane.txt}
file).
\item To change the color scale, go to \textsf{Display Tab} $\rightarrow$
\textsf{Color} $\rightarrow$ \textsf{Edit Color Map} and reselect
lower and upper limits, or change the color scheme.
\item Now load in the source and receiver location file by \textsf{File}
$\rightarrow$ \textsf{Open data}, select \texttt{sr.vt}k, and click
the \textsf{Accept} button. Choose \textsf{Filter} $\rightarrow$
\textsf{Glyph}, and represent the points by `\textsf{spheres}'.
\item For more information about ParaView, see the ParaView Users Guide \urlwithparentheses{www.paraview.org/files/v1.6/ParaViewUsersGuide.PDF}.
\end{enumerate}
\end{enumerate}
%
\begin{figure}[H]
\noindent \begin{centering}
\includegraphics[scale=0.6]{figures/3D-S-Kernel}
\par\end{centering}

@@

\begin{enumerate}
\item First the script directs the scheduler to save its own output and
output from \texttt{stdout} into ~\\
\texttt{\small in\_out\_files/OUTPUT\_FILES/\%J.o},
where \texttt{\%J} is short-hand for the job-ID; it also tells the
scheduler what version of \texttt{mpich} to use (\texttt{mpich\_gm})
and how to name this job (\texttt{go\_mesher\_solver\_lsf}).
\item The script then creates a list of the nodes allocated to this job
by echoing the value of a dynamically set environment variable \texttt{LSB\_MCPU\_HOSTS}
and parsing the output into a one-column list using the Perl script
\texttt{utils/Cluster/lsf/remap\_lsf\_machines.pl}. It then creates a set of scratch
directories on these nodes (\texttt{\small /scratch/}~\\
\texttt{\small \$USER/DATABASES\_MPI}) to be used as the \texttt{LOCAL\_PATH}
for temporary storage of the database files. The scratch directories
are created using \texttt{shmux}, a shell multiplexor that can execute
the same commands on many hosts in parallel. \texttt{shmux} is available
from Shmux \urlwithparentheses{web.taranis.org/shmux/}. Make sure that the \texttt{LOCAL\_PATH}
parameter in \texttt{in\_data\_files/Par\_file} is also set properly.
\item The next portion of the script launches the mesher, database generator and then the solver
using \texttt{run\_lsf.bash}.
\item The final portion of the script collects the seismograms and performs
clean up on the nodes, using the Perl scripts \texttt{collect\_seismo\_lsf\_multi.pl}
and \texttt{cleanmulti.pl}.
\end{enumerate}
\begin{lyxcode}
\#!/bin/bash~-v~

@@

rotate.pl~-l~0~-L~180~SEM/{*}.BXE.semd.sac.bp~
\end{lyxcode}
where the first command performs rotation on the SAC data obtained
through Seismogram Transfer Program (STP) \urlwithparentheses{http://www.data.scec.org/STP/stp.html},
while the second command rotates the processed SEM synthetics.

@@

To report bugs or suggest improvements to the code, please send an
e-mail to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}
or Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu}, and/or use our online
bug tracking system Roundup \urlwithparentheses{www.geodynamics.org/roundup}.

@@

OpenDX \urlwithparentheses{http://www.opendx.org} is open-source based on IBM Data
Explorer, AVS \urlwithparentheses{http://www.avs.com} is a trademark of Advanced
Visualization Systems, and ParaView \urlwithparentheses{http://www.paraview.org}
is an open-source visualization platform.

Please e-mail your feedback, questions, comments, and suggestions
to Jeroen Tromp \urlwithparentheses{jtromp-AT-princeton.edu} or to the CIG Computational Seismology Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.
